Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-10-04 arxiv/diff_cov.tex
revise
finish section 4

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	Assume further that: a) the singular values of $M$ are all greater than $\tau$ and less than $1/\tau$; b) the sample size of task two is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:

	\noindent (i) The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Our result extends a well-known result on the inverse of sample covariance matrices.
In particular, by setting $X_1 = 0$, we obtain the second result in Fact \ref{lem_minv}.
When $X_1=0$, it is not hard to solve $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ from equation \eqref{eq_a12extra}, and our result reduces to the case of a single sample covariance matrix.
For a multivariate Gaussian random matrix, this result follows from the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For a non-Gaussian random matrix, this result can be obtained using the well-known Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).

As mentioned in the beginning of this section, different tasks may have different sample sizes and different feature covariance matrices.
Under these scenarios, Theorem \ref{thm_main_RMT} shows the bias and variance asymptotics for the two-task case.
Next, we use this result to study the impact of varying sample sizes and covariate shifts in hard parameter sharing.

\begin{example}[Sample ratio]\label{ex_sample_ratio}
	Our first example considers the impact of varying sample sizes on the prediction loss of hard parameter sharing.
	We illustrate the above result in the random-effects model described in Section \ref{sec_hps}.
	We further assume that the population covariance matrices of both tasks are isotropic.

	To apply Theorem \ref{thm_main_RMT}, we solve the self-consistent equations \eqref{eq_a12extra} and \eqref{eq_a34extra} by using $\lambda_i = 1$ for all $1\le i\le p$.
	For the variance limit, since $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2 - p}$, we have that $\tr[(a_1\Sigma^{(1)} + a_2\Sigma^{(2)})^{-1}] / (n_1 + n_2) = \frac{p}{n_1 + n_2 - p}$.
	For the bias limit, we first solve $a_1$ and $a_2$ from equation \eqref{eq_a12extra}, and then solve $a_3$ and $a_4$ by using the solutions to equation \eqref{eq_a34extra}.
	Using the solved weights in the bias and variance limit equations, we obtain the following result.

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) task one's sample size is at least $n_1 \ge \min(\frac{17}{\varepsilon^2}, 50) \cdot p$ for a fixed positive value $\varepsilon$, and task two's sample size is at least $n_2 \ge \frac{3}{2}p$;
	(ii) the noise variance is smaller than the signal strength: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
	(iii) the task-specific variance of $\beta_i$ is much smaller than the signal strength: $d^2 \le p^{-c_{\varphi}}{\kappa^2}$. \HZ{check}
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} + p^{-c_{\varphi}} \cdot \OO(d^2 + \sigma^2).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
%	 for a large constant $C>0$.
\end{corollary}

@@

In the above result, we see that the prediction loss of hard parameter sharing consists of a bias term that scales with task-specific variance and a noise variance term.
We use the above result to explain an interesting phenomenon, where adding task one's samples helps task two initially, but hurts eventually.
In particular, consider the prediction loss of hard parameter sharing as we fix $n_2$ and vary $n_1$.
First, the noise variance term always reduces as $n_1$ increases.
Second, when $n_2 \ge \frac{3}{2} p$, the bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
Therefore, we conclude that the \textit{optimal} sample ratio between $n_1$ and $n_2$ is one of three cases:
(i) $n_1 = 0$, and the prediction loss reduces to that of single-task learning for task two;
(ii) $n_1 = \infty$, and the prediction loss reduces to the task-specific variance $d^2$;
(iii) the critial point where the derivative of the bias and variance terms is zero.
Furthermore, we observe the following trend by comparing the conditions of $n_1 = 0$ and $n_1 = \infty$:
\begin{itemize}
	\item If $d^2$ is smaller than the prediction loss of STL for task two, then the prediction loss of hard parameter sharing is always lower than STL. That is, task one always transfers positively to task two.
	\item Otherwise, as the sample ratio between $n_1$ and $n_2$ increases, task one transfers positively to task two initially, but transfers negatively eventually.
\end{itemize}
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we observe negative transfer.
%of MTL increases and converges to $pd^2$ as the sample ratio $\rho_1/\rho_2$ increases, and the variance reduction $\frac{\sigma^2}{ \rho_2-1}  -\frac{\sigma^2}{\rho_1+\rho_2-1}$ also increases and converges to $\frac{\sigma^2}{\rho_2-1}$ as $\rho_1/\rho_2$ increases. The tradeoff between these two terms depends on $\rho_1$ in a nonlinear way that is more complicated than the one in the case of varying model distance. But from our results, one can check the following phenomena. If $\frac{\sigma^2}{\rho_2-1} > pd^2$, then the transfer is always positive. On the other hand, if $\frac{\sigma^2}{\rho_2-1} < pd^2$, then there is a transition from the positive transfer to negative transfer as $\rho_1/\rho_2$ increases.
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we will observe negative transfer.
%Figure \ref{fig_size} provides a simulation result for such a setting (described in Appendix \ref{app_synthetic}).
%We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
%There are several unexplained observations near $y = 0$ caused by the concentration error on the right-hand side of \eqref{cor_MTL_error}.
\end{example}

@@

\begin{example}[Covariate shift]\label{sec_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
In our final example, we consider what happens when the covariance matrices are also different between different tasks, which are often called ``covariate shift''.
Our results apply to arbitrary ``covariate shift'' and we consider a simple setting of two tasks.
As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
To illustrate our result, we assume that the first half of the singular values of $\Sigma^{1/2}\Sigma^{-1/2}$ are all equal to a parameter $\lambda$ and the rest are all equal to one.
Our results imply that the prediction loss of the hard parameter sharing estimator is given as follows.

	In the simulation of Figure \ref{fig_covariate}, we observe the following two phases as  $n_1 / p$ increases.
When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
We provide a theoretical justification now.

\begin{proof}[Theoretical justification of Example \ref{ex_complement}]
We denote the test error as $\te(\hat \beta_{2}^{\MTL},\lambda)$ in the setting where $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$. Then equations in equation \eqref{eq_a12extra} become
\be\label{compleeq} a_1 + a_2 = 1 - \frac{1}{\rho_1 + \rho_2},  \ \ a_1 + \frac{1}{2(\rho_1 + \rho_2)}\cdot \bigbrace{\frac{a_1}{a_1 + \lambda^2 a_2} + \frac{a_1}{a_1 + \lambda^{-2} a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \ee
%It's not hard to verify that there is only one valid solution $(a_1,a_2)$ to equation \eqref{compleeq}.
After solving these, with equation \eqref{gvar_extra} we get that w.h.p.,
\be\label{testcomple}
 \te(\hat \beta_{2}^{\MTL},\lambda)= \frac{\sigma^2}{2(\rho_1 + \rho_2)}(1+\OO(p^{-\e}))\cdot f(\lambda) ,\quad f(\lambda): = \frac{1}{\lambda^{-2}{a_1} + a_2} + \frac{1}{\lambda^2a_1 + a_2}.\ee

%First we notice that the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$. In fact, if $n_1=n_2$, then it is easy to observe that $a_1=a_2=(1-\gamma)/2$ is the solution to equation \eqref{compleeq}, where we denote $ \gamma=p/(n_1+n_2)$. Then for any $\lambda$, the test error in equation \eqref{testcomple} takes the value
%$$te(\lambda)= \frac{\gamma}{2}\frac{1}{(1-\gamma)/2}=\frac{p}{n_1+n_2-p}.$$

%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
We now study the behavior of $f$ as $\lambda$ changes.
%This phenomenon can be also explained using our theory.
We abbreviate $\gamma:=(\rho_1 + \rho_2)^{-1}$. Then with equation \eqref{compleeq}, we can rewrite
$$f(\lambda)= \frac{1}{\lambda^{-2}{a_1} + (1-\gamma - a_1)} + \frac{1}{\lambda^2a_1 + (1-\gamma - a_1)}.$$
Then we can compute that
\begin{align*}
f(\lambda) - f(1)&= \frac{ \lambda^2-1}{1-\gamma} a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{(\lambda^2-1)^2}{1-\gamma}  a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
\end{align*}
From this expressions, we observe the following behaviors.
\begin{itemize}
\item[(i)] If $n_1>n_2$, we have $a_1>(1-\gamma)/2$ (because $a_1>a_2$ as observed from the equation \eqref{compleeq}). Hence $f(\lambda)>f(1)$, which gives $\te(\hat \beta_{2}^{\MTL},\lambda)>\te(\hat \beta_{2}^{\MTL},1)$.

\item[(ii)] If $n_1< n_2$, we have $a_1< (1-\gamma)/2$. Hence $f(\lambda)< f(1)$, which gives $\te(\hat \beta_{2}^{\MTL},\lambda)<\te(\hat \beta_{2}^{\MTL},1)$.

\item[(iii)] If $n_1=n_2$, we have $f(\lambda)=f(1)=2/(1-\gamma)$, which means $\te(\hat \beta_{2}^{\MTL},\lambda)$ and $\te(\hat \beta_{2}^{\MTL},1)$ are roughly the same. %, which explains why the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$.
\end{itemize}
This also partially justifies Proposition \ref{prop_covariate}. %Theorem  the observations in Figure \ref{fig_model_shift_phasetrans} (c),
\end{proof}

%We compare two cases: (i) when $M = \id_{p\times p}$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
%Hence, $\lambda$ measures the severity of the covariate shift.
%	Figure \ref{fig_covariate} shows a simulation of this setting by varying $\lambda$.
%	We observe that as source/target sample ratio increases, the performance gap between the two cases increases.
%	\end{example}

>>>>

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	Assume that: a) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$; b) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:

	\noindent (i) The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Our result extends a well-known result on the inverse of sample covariance matrices.
If $X_1$ is zero, then we obtain Fact \ref{lem_minv}(ii) as a special case.
To see this, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to the variance limit.
\HZ{go to intro}
For a multivariate Gaussian random matrix, this result follows from the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For a non-Gaussian random matrix, this result can be obtained using the well-known Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).

%As mentioned in the beginning of this section, different tasks may have different sample sizes and different feature covariance matrices.
%Under these scenarios, Theorem \ref{thm_main_RMT} shows the bias and variance asymptotics for the two-task case.
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we use our result to study how varying sample sizes and covariate shifts affects hard parameter sharing, respectively.

\begin{example}[Sample ratio]\label{ex_sample_ratio}
	We first consider the impact of varying sample sizes.
	Consider the random-effects model from Section \ref{sec_hps}, with both tasks having an isotropic population covariance matrix.

	In applying Theorem \ref{thm_main_RMT} to the above setting, we first solve the self-consistent equations \eqref{eq_a12extra} by using $\lambda_i = 1$ for all $1\le i\le p$.
	This gives us the variance limit $\tr[(a_1\Sigma^{(1)} + a_2\Sigma^{(2)})^{-1}] / (n_1 + n_2) = \frac{p}{n_1 + n_2 - p}$, since we have that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2 - p}$.
	Secondly, we solve the self-consistent equations \eqref{eq_a34extra} given that we already know $a_1$ and $a_2$.
	This gives us the bias limit.
	Combined together, we obtain the following corollary, which requires natural assumptions over the random-effects model.

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) task one's sample size is at least $n_1 \ge \min(\frac{17}{\varepsilon^2}, 50) \cdot p$ for a fixed positive value $\varepsilon$, and task two's sample size is at least $n_2 \ge \frac{3}{2}p$;
	(ii) the noise variance is smaller than the signal strength: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) the task-specific variance of $\beta_i$ is much smaller than the shared signal strength: $d^2 \le p^{-c_{\varphi}}{\kappa^2}$.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} + p^{-c_{\varphi}} \cdot \OO(d^2 + \sigma^2).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
%	 for a large constant $C>0$.
\end{corollary}

@@

Our result shows that the prediction loss of hard parameter sharing consists of a bias term that scales with task-specific variance $d^2$ and a noise variance term that scales with $\sigma^2$.
In the following, we explain an interesting phenomenon, where adding task one's samples helps task two initially, but hurts eventually.
Consider the prediction loss if we fix $n_2$ and increase $n_1$.
First, the noise variance term in Corollary \ref{cor_MTL_loss} always reduces as $n_1$ increases.
Second, the bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
Combined together, we observe that the \textit{optimal} sample ratio between $n_1$ and $n_2$ should be among one of following three cases:
(i) $n_1 = 0$, in which case the prediction loss reduces to $\frac{\sigma^2 p}{n_2 - p}$;
(ii) $n_1 = \infty$, in which case the prediction loss reduces to $d^2$;
(iii) the critial point where the derivative of the bias and the variance terms w.r.t. $n_1$ equals zero.

Based on this observation, we have the following trend when we fix $n_2$ and increase $n_1$:
\begin{itemize}
	\item If $d^2$ is smaller than the prediction loss of STL for task two $\frac{\sigma^2 p}{n_2 - p}$, then the prediction loss of hard parameter sharing is always lower than STL. That is, task one always transfers positively to task two.
	\item Otherwise, task one transfers positively to task two initially, but transfers negatively eventually.
\end{itemize}
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we observe negative transfer.
%of MTL increases and converges to $pd^2$ as the sample ratio $\rho_1/\rho_2$ increases, and the variance reduction $\frac{\sigma^2}{ \rho_2-1}  -\frac{\sigma^2}{\rho_1+\rho_2-1}$ also increases and converges to $\frac{\sigma^2}{\rho_2-1}$ as $\rho_1/\rho_2$ increases. The tradeoff between these two terms depends on $\rho_1$ in a nonlinear way that is more complicated than the one in the case of varying model distance. But from our results, one can check the following phenomena. If $\frac{\sigma^2}{\rho_2-1} > pd^2$, then the transfer is always positive. On the other hand, if $\frac{\sigma^2}{\rho_2-1} < pd^2$, then there is a transition from the positive transfer to negative transfer as $\rho_1/\rho_2$ increases.
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we will observe negative transfer.
%Figure \ref{fig_size} provides a simulation result for such a setting (described in Appendix \ref{app_synthetic}).
%We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
%There are several unexplained observations near $y = 0$ caused by the concentration error on the right-hand side of \eqref{cor_MTL_error}.
\end{example}

@@

\begin{example}[Covariate shift]\label{sec_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example studies the impact of varying covariate shifts on the variance limit.
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Recall that the matrix $M$ captures the geometry of covariate shifts.
Suppose that half of the singular values of $M$ are equal to a parameter $\lambda > 0$ and the other half are equal to $\lambda^{-1}$.
We observe the following dichotomy on the impact of covariate shifts.
\begin{itemize}
	\item If $n_1 \ge n_2$, then the variance (limit) of hard parameter sharing is smallest when $\lambda = 1$ or there is no covariate shift.
	\item If $n_1 < n_2$, then having covariate shifts ($\lambda \neq 1$) always helps reduce variance (limit).
\end{itemize}
To see this, we first write the variance limit for different values of $\lambda$ as follows
\begin{align*}
	\bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}
	= \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}
	= \frac{1}{2}\bigbrace{(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}.
\end{align*}
Denote the above function by $f(\lambda)$.
We compare $f(\lambda)$ with $f(1)$, and observe that
\[ 2f(\lambda) - 2f(1) = {(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1})^{-1}
	+ (\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1)^{-1}  - \frac{2(n_1 + n_2)}{n_1 + n_2 - p}, \]
where we used the first part of equation \eqref{lem_cov_shift_eq}.
Next, one can verify that the above equation is equal to $(2a_1 - \frac{n_1 + n_2 - p} {n_1 + n_2}) g(\lambda, a_1)$, for a fixed and positive function $g(\lambda, a_1)$ (details omitted).
Hence, whether or not $f(\lambda)$ is small than $f(1)$ is decided by $a_1$.
To solve $a_1$, we present an informal argument that $a_1 \ge \frac{(n_1 + n_2 - p)}{2(n_1 + n_2)}$ if and only if $n_1 \ge n_2$.
Intuitively, with the current spectrum of $M$, $a_1$ and $a_2$ have a symmetric role in the self-consistent equations.
Hence, when $n_1 \ge n_2$, we expect that $a_1 \ge a_2$, hence $a_1 \ge \frac{1}{2}(1 - \frac{p}{n_1 + n_2})$.
Vice serva when $n_1 < n_2$.
A formal proof requires solving the self-consistent equations \eqref{lem_cov_shift_eq} by replacing half of the singular with $\lambda$ and the other half with $\lambda^{-1}$.
We omit the details.
%We denote the test error as $\te(\hat \beta_{2}^{\MTL},\lambda)$ in the setting where $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$. Then equations in equation \eqref{eq_a12extra} become
%\begin{align*}
%	a_1 + \frac{p}{2(n_1 + n_2)}\cdot \bigbrace{\frac{a_1}{a_1 + \lambda^2 a_2} + \frac{a_1}{a_1 + \lambda^{-2} a_2}} = \frac{n_1}{n_1 + n_2}.
%\end{align*}
%It's not hard to verify that there is only one valid solution $(a_1,a_2)$ to equation \eqref{compleeq}.
%After solving these, with equation \eqref{gvar_extra} we get that w.h.p.,
%\be\label{testcomple}
% \te(\hat \beta_{2}^{\MTL},\lambda)= \frac{\sigma^2}{2(\rho_1 + \rho_2)}(1+\OO(p^{-\e}))\cdot f(\lambda) ,\quad f(\lambda): = \frac{1}{\lambda^{-2}{a_1} + a_2} + \frac{1}{\lambda^2a_1 + a_2}.\ee
%First we notice that the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$. In fact, if $n_1=n_2$, then it is easy to observe that $a_1=a_2=(1-\gamma)/2$ is the solution to equation \eqref{compleeq}, where we denote $ \gamma=p/(n_1+n_2)$. Then for any $\lambda$, the test error in equation \eqref{testcomple} takes the value
%$$te(\lambda)= \frac{\gamma}{2}\frac{1}{(1-\gamma)/2}=\frac{p}{n_1+n_2-p}.$$
%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
%We now study the behavior of $f$ as $\lambda$ changes.
%This phenomenon can be also explained using our theory.
%We abbreviate $\gamma:=(\rho_1 + \rho_2)^{-1}$. Then with equation \eqref{compleeq}, we can rewrite
%$$f(\lambda)= \frac{1}{\lambda^{-2}{a_1} + (1-\gamma - a_1)} + \frac{1}{\lambda^2a_1 + (1-\gamma - a_1)}.$$
%Then we can compute that
%\begin{align*}
%f(\lambda) - f(1)&= \frac{ \lambda^2-1}{1-\gamma} a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
%&= \frac{(\lambda^2-1)^2}{1-\gamma}  a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
%\end{align*}
Thus, we conclude that if $n_1 \ge n_2$, then $f(\lambda) > f(1)$; if $n_1< n_2$, then $f(\lambda)< f(1)$. % which gives $\te(\hat \beta_{2}^{\MTL},\lambda)<\te(\hat \beta_{2}^{\MTL},1)$.
%\item[(iii)] If $n_1=n_2$, we have $f(\lambda)=f(1)=2/(1-\gamma)$, which means $\te(\hat \beta_{2}^{\MTL},\lambda)$ and $\te(\hat \beta_{2}^{\MTL},1)$ are roughly the same. %, which explains why the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$.
%\end{itemize}
%This also partially justifies Proposition \ref{prop_covariate}. %Theorem  the observations in Figure \ref{fig_model_shift_phasetrans} (c),

%We compare two cases: (i) when $M = \id_{p\times p}$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
%Hence, $\lambda$ measures the severity of the covariate shift.
%	Figure \ref{fig_covariate} shows a simulation of this setting by varying $\lambda$.
%	We observe that as source/target sample ratio increases, the performance gap between the two cases increases.
